---
title: "Data Tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# notes: YOU CAN MAKE YOUR PARENTHESIS RAINBOW BY GOING GLOBAL OPTIONS-> CODE-> DISPLAY->RAINBOW PARENTHESIS

# Library Calls
```{r}
library(dplyr)
library(readr)
library(tidyr)
```

# Data Sources
- read in the data
```{r}

catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1",show_col_types = TRUE)
```


# Clean and Reshape data
- remove unnecessary columns
- check column typing
- reshape data

Remove columns for all and notes
```{r}
catch_data <- catch_original %>% 
  select(-All, -notesRegCode)

head(catch_data)

glimpse(catch_data)
```

Coerce Chinook column into a numeric instead of a character
```{r}
catch_clean <-catch_original%>%
  mutate(Chinook= as.numeric(Chinook))

```

Locating the row that's throwing an NA value
```{r}
i <- which(is.na(catch_clean$Chinook))

i

catch_original[i,]

```

Replacing the value that's creating an NA
```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.integer(Chinook))

head(catch_clean)
summary(catch_clean)
```

Reshaping the data longer
```{r}
catch_long<- catch_clean%>%
  pivot_longer(cols=-c(Region,Year),names_to = "Species",values_to="Catch" )

head(catch_long)
```

Reshaping the data wider
```{r}
catch_wide<-catch_long %>%
  pivot_wider(names_from = "Region",values_from = "Catch")


```

```{r}
catch_long<-catch_long %>%
  rename(Catch_thousands = Catch)%>%
  mutate(Catch = Catch_thousands * 1000)%>%
  select(-Catch_thousands)

head(catch_long)
```


# Split -apply -combine
```{r}
mean_region<-catch_long %>%
  group_by(Region)%>%
  summarise(catch_mean=mean(Catch),
            n = n())

mean_region

mean_year<-catch_long%>%
  group_by(Year)%>%
  summarise(catch_mean=mean(Catch),
            n = n())

mean_year

mean_spp<-catch_long%>%
  group_by(Species)%>%
  summarise(catch_mean=mean(Catch),
            n = n())
mean_spp

var<-catch_long%>%
  filter(Year > 1899)%>%
  group_by(Species)%>%
  summarise(catch_mean=mean(Catch), .groups="drop")%>%
  arrange(-catch_mean)

var
```

# Join to Regions dataset
```{r}
```

